---
output: github_document
title:  Lecture 5 (ggplot2) - Summary
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
Here are some of the key points from the lecture. See also the sample code solutions in the code folder.

* To show the workings of ggplot2, we will use the mpg dataset, and the tibble library.

```{r}
library(ggplot2)
library(tibble)

glimpse(mpg)

```

* ggplot2 works in layers, with data and mapping. The data must be in a tibble/data frame. Mappings involved the aes() function and map variables onto plot features. A simple example is shown below.

```{r}
ggplot(data=mpg) + geom_point(aes(x=displ,y=cty))+ylab("City MPG")+xlab("Displacement")
```

* Point colours can be set directly as part of geom_point. We can also layer plots on one another, for example, showing the different between city MPG and highway MPG.

```{r}
ggplot(data=mpg) + geom_point(aes(x=displ,y=cty), colour="red")+
  geom_point(aes(x=displ,y=hwy), colour="blue")+
  ylab("MPG")+xlab("Displacement")
```

* The aes() function can also be used to colour points (automatically) by category. Note that the colour parameter is inside the aes() function. This is important!

```{r}
ggplot(data=mpg) + geom_point(aes(x=displ,y=cty,colour=class))
```

* Facets can alse be used to explore data, by splitting the graphs out into new categories
```{r}
ggplot(data=mpg) + geom_point(aes(x=displ,y=cty,colour=class))+facet_wrap(~manufacturer)
```

* Facets can alse be used in a grid fashion for two categorical variables.
```{r}
ggplot(data=mpg) + geom_point(aes(x=displ,y=cty,colour=manufacturer))+facet_grid(cyl~class)
```

* Compare the previous plot with the following command
```{r}
table(mpg$cyl,mpg$class)
```

* Statistical models can also be layered on a graph
```{r}
ggplot(data=mpg, aes(x=displ,y=cty))+geom_point()+geom_smooth(method = "lm")
```

* To show further plots, the dataset ggplot2::diamonds is used

```{r}
glimpse(diamonds)
```

* Bar charts can process observations to produce overall numeric counts
```{r}
ggplot(data=diamonds, aes(x=cut))+geom_bar()
```

* These can be viewed by categories (e.g. using fill=clarity)
```{r}
ggplot(data=diamonds, aes(x=cut,fill=clarity))+geom_bar()
```

* Boxplots can also be drawn
```{r}
ggplot(data=diamonds, aes(x=cut,y=carat))+geom_boxplot()
```

**Additional Examples**

* [Top 50 ggplot2 Visualizations](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)
* [Exploration of MPG Dataset by Shailesh Kumar](https://rpubs.com/shailesh/mpg-exploration)
* [R Cookbook by Winston Chang](http://www.cookbook-r.com/Graphs/)

