---
title: '4. Base R- Data Frames'
author: "CT5102 - J. Duggan"
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "seahorse"
    fonttheme: "structurebold"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(dplyr)
```


## Recap - R Data Types
 
|| Homogenous| Heterogenous |
|:------|:----------|:-----------|
|1d | Atomic Vector|List|
|2d | Matrix|**Data Frame/Tibble**|
|nd | Array| |

- The most common way of storing data in R
- Under the hood, a data frame is a list of equal-length vectors
- A two-dimensional structure, it shares properties of both a list and a matrix



## Creating a data frame

```{r,echo=T}
d <- data.frame(x=1:3, y = LETTERS[1:3], z = letters[1:3])
d
d$x
d$y
d$z
```

## summary function with data frames

```{r,echo=T}
d <- data.frame(x=1:3, y = LETTERS[1:3], 
                z = letters[1:3])
d
summary(d)
```

## mtcars data frame
A data frame with 32 observations on 11 variables.

- **mpg**	Miles/(US) gallon	
- **cyl**	Number of cylinders	
- **disp**	Displacement (cu.in.)	
- **hp**	Gross horsepower	
- **drat**	Rear axle ratio	
- **wt**	Weight (1000 lbs)	
- **qsec**	1/4 mile time	
- **vs**	V/S	
- **am**	Transmission (0 = automatic, 1 = manual)	
- **gear**	Number of forward gears	
- **carb**	Number of carburetors

## mtcars sample data

```{r,echo=T}
knitr::kable(mtcars[1:10,1:6])
```

## mtcars using **str()**
```{r,echo=T}
str(mtcars)
```
## head() and tail() functions
```{r,echo=T}
head(mtcars[,1:6])
tail(mtcars[,1:6])
```

## Subsetting rows
```{r,echo=T}
mtcars[mtcars$gear == 5,]
```

## Accessing rows/columns
```{r,echo=T}
mtcars[1:10,1:6]
```

## Filtering rows and columns
```{r,echo=T}
mtcars[mtcars$cyl == 6,c("mpg","cyl")]
```

## Challenge 4.1
- List all the cars that have an **mpg** greater than the average
- List the car(s) with the greatest displacement (**disp**)

## subset()
- The subset( ) function can also be used to select variables and observations.
- Takes the data frame, the conditions, and the columns to return
```{r,echo=T}
subset(mtcars,cyl==6,select=c("mpg","cyl"))

```

## Processing a data frame with functions - sapply()
- As a data frame is also a list, the apply family of functions can be used to process the list
- sapply and lapply process a data frame on a column by column basis
```{r,echo=T}
sapply(subset(mtcars,cyl==6,select=c("mpg","cyl")),
       function(col){min(col)})
```

## Processing a data frame with functions - apply()
- The apply function can be used to process the data frame by row or by column
```{r,echo=T}
apply(subset(mtcars,cyl==6,select=c("mpg","cyl")),2,
       function(col){min(col)})
apply(subset(mtcars,cyl==6,select=c("mpg","cyl")),1,
       function(row){max(row)})
```

## Adding new columns to a data frame
- Often the initial data set may not contain sufficient information for analysis
- Adding new variables (columns) is an important feature to have
- Data frames support this: columns can be combined or new information used

```{r,echo=T}
mtcars$name <- rownames(mtcars)
mtcars[1:5,-(1:8)]
```

## Deleting a column
- To delete a column, set it to NULL

```{r,echo=T}
mytest <- mtcars
mytest[1,1:3]
mytest$mpg <- NULL
mytest[1,1:3]
```
## Challenge 4.2
Create a new column on mtcars that contains kilometers per gallon. 

## Missing data - complete.cases()

```{r,echo=T}
d <- data.frame(x=1:3, y = LETTERS[1:3], 
                z = letters[1:3])
d[2,3] <- NA
d
complete.cases(d)
d[complete.cases(d),]
```

## Exploring complete.cases()
```{r,echo=T}
mtcars1 <- subset(mtcars,select = mpg:disp)
sum(complete.cases(mtcars1))

mtcars1[4,3] <- NA
mtcars1[4,]
mtcars1[!complete.cases(mtcars1),]
```

## Challenge 4.3
- Explore the data set nycflights13::flights
- Find the total number of incomplete cases
- Use sapply() to find the number of missing values for each column.

## The tibble
- Tibbles are data frames, but they tweak some older behaviours to make life a little easier
- One of the unifying features of the tidyverse
- To coerce a data frame to a tibble, use as_tibble()
- A tibble can be created from individual vectors using tibble()


```{r,echo=T}
t <- tibble(x=1:3, y = LETTERS[1:3], z = letters[1:3])
t
```

## Tibble abbreviations

```{r,echo=T}
t
```

|Abbreviation| Data Type |
|:------|:--------|
|int | integers |
|dbl | double (numeric)|
|chr | character vectors| 
|dttm | date-times| 
|fctr | categorical| 
|date | dates |


## Summary - Data Frames
- Data frames/tibbles are the most common way of storing heterogeneous data in R
- Under the hood, a data frame is a list of equal-length vectors, and shares properties of both a list and a matrix
- Key for processing rectangular data, ideally in "tidy" format (every row is an observation, every column a variable)

## Summary - Base R
|| Homogenous| Heterogenous |
|:------|:----------|:-----------|
|1d | Atomic Vector|List|
|2d | *Matrix*|Data Frame/Tibble|
|nd | *Array*| |

- Atomic Vectors
- Lists
- Functions and Functionals
- Data Frames & Tibbles



```{r,echo=F,fig.width=2, fig.height=2}
# this is a hack to call plot but make the plot tiny
plot(1:1,axes=FALSE,xlab = "", ylab="",cex = .001)
```


