---
title: "Exploring Heights"
date: '`r lubridate::today()`'
output:
  html_document: default
  pdf_document: default
  word_document: default
bibliography: ref.bib
---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)

path <- "/Users/jim/Dropbox/R Projects/CT5102/code/course/12 Summary/01 R Markdown/"
d <- read_csv(paste0(path,"HopeCollegeHeights.csv"))
d <- d %>% mutate(Gender=as.factor(ifelse(Gender==1,"Female","Male")))
```

###Introduction
This document summarises provides an initial analysis of height data from the 1990s, as sourced from [@Link1].

###Summary of the Data
The data has `r nrow(d)` observations, with the following summary table:
```{r, echo=FALSE}
sd <- d %>% group_by(Gender) %>% 
              summarise(Mean=mean(Height),
                        Median=median(Height),
                        SDev=sd(Height),
                        Number=n()
                        )
knitr::kable(
  sd,
  caption="Summary of Heights Data"
)
```

###Visualisation
The first graphs shows the two cohorts.

```{r,echo=FALSE}
ggplot(d, aes(Height)) + 
  geom_histogram(binwidth = 1) + 
  facet_grid(Gender ~ .)
```

We then show the two cohorts in a single graph.

```{r, echo=FALSE}
ggplot(d, aes(Height, fill = Gender)) + 
  geom_histogram(binwidth = 1)
```

###Statistical Test
Finally, a student-t test is performed on the two data sets, confirming the statistical difference.

```{r, echo=FALSE}
males   <- d %>% filter(Gender=="Male")   %>% select(Height) %>% pull()

females <- d %>% filter(Gender=="Female") %>% select(Height) %>%pull()

t <- t.test(males,females)

t
```


###References