---
output: github_document
title:  Lecture 3 (Matrices and Functions) - Summary
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
Here are some of the key points from the lecture. See also the sample code solutions in the code folder.

* The ... operator is useful for capturing additional parameters that are sent to a function. These can also be forwared to other functions.

```{r}
# The ... operator can be examined in a function by using it to create a list
test_dot <- function(...){
  list(...)
}

test_dot(1:2, 3:4, c(T,F))
```

```{r}
# The ... operator can be used to capture additional parameters and forward these to other functions
mysum <- function(v,...){
  sum(v,...)
}

x <- c(1,2,3,4, NA)

mysum(x)

mysum(x,na.rm=T)
```

* do.call() can be used to call a function with arguments that are specified in a list


```{r}
args <- list(c(1:10,NA), na.rm = TRUE)

ans <- do.call(sum,args)
```

* Replacement functions can be used to modify the argument. Examples in R include names(), diag(), and dim()

```{r}
v <- 1:12
v

dim(v) <- c(3,4)
v
```

* Replacement functions can also be written.
```{r}
`second<-` <- function(y, value) {
  x[2] <- value
  x
}
x <- 1:10
second(x) <- 500L
x
```

* Matrices can be created from vectors, and are filled in column order (by default)
```{r}
v <- 1:12
v

m <- matrix(v,nrow = 3,ncol=4)
m
```

* Matrices can have rows and columns named
```{r}
v <- 1:12
v

m <- matrix(v,nrow = 3,ncol=4)
rownames(m) <- letters[1:3]
colnames(m) <- LETTERS[1:4]
m
```

* Matrices can be subsetted by row or by column. Leaving a blank will display either all rows or all columns

```{r}
m
# show first two rows and all columns
m[1:2,]
```

```{r}
m
# show first two columns and all rows
m[,1:2]
```

* Matrices can be subsetted by row or column name

```{r}
m
# show first two columns and all rows
m[,c("B","C")]
```

* A functional is a function that takes a function as an input and returns a vector as output. Commonly used in R to iterate over data. The apply family of functions are very useful, for example, lapply(x,f,args)

```{r}
l <- list(1:2, 3:4, 5:6)
lapply(l,max)
sapply(l,max)
```

* Anonymous functions can also be written
```{r}
x <- seq(-10,10,by = 0.1)

a <- 3
b <- -10
c <- 5

y1 <- lapply(x, function(x){
  a*x^2+b*x+c
})

plot(x,y1)

```

* The apply function can be used to process matrices on a row by row (paramter=1), or a column by column basis (parameter =2)

```{r}
set.seed(10)
N=10

cs1 <- rnorm(N,72,10)
cs2 <- rnorm(N,65,7)
cs3 <- rnorm(N,80,9)
cs4 <- rnorm(N,55,7)
cs5 <- rnorm(N,61,5)

m   <-cbind(cs1, cs2, cs3, cs4, cs5)

m <- cbind(m, avr=apply(m,1,mean))
m
```
