---
output: github_document
title:  Lecture 7 (Relational Data and tidyr) - Summary
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
The package **dplyr** also supports mutating and filtering joins. For the following two tables, we can explore different functions to join the tables.

```{r}
library(dplyr)
tab_x <- tibble(key=c(1,2,3),
                val_x=c("x1","x2","x3"))

tab_x

tab_y <- tibble(key=c(1,2,4),
                val_y=c("y1","y2","y3"))

tab_y

```

* **inner_join()** matches pairs of observations when their keys are equal. Unmatched rows are not included in the result.

```{r}
inner_join(tab_x,tab_y)
```

* **left_join()** keeps all observations in x
```{r}
left_join(tab_x,tab_y)
```

* **right_join()** keeps all observations in y

```{r}
right_join(tab_x,tab_y)
```

* **full_join()** keeps all observations in x and y

```{r}
full_join(tab_x,tab_y)
```

* **semi_join()**  (filtering join) keeps all observations in x that have a match in y
```{r}
semi_join(tab_x,tab_y)
```


* **anti_join()**  (filtering join) drops all observations in x that have a match in y
```{r}
anti_join(tab_x,tab_y)
```