---
output: github_document
title:  Lecture 4 (Data Frames) - Summary
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
Here are some of the key points from the lecture. See also the sample code solutions in the code folder.

* A data frame stores heterogeneous data, and the package datasets contains good examples (i.e. mtcars - 32 observations of 11 variables)

```{r}
str(mtcars)

head(mtcars)
```

* Matrix notation can be used to access data frames

```{r}
mtcars[1:4,]
mtcars[1:10,c("mpg","cyl","disp")]

```

* The list $ operator can be used to access column data, and perform conditional operations
```{r}
mtcars$mpg
mtcars$mpg > mean(mtcars$mpg)
```

* Combining these can be useful for subsetting
```{r}
mtcars[mtcars$mpg > mean(mtcars$mpg),]
```

* New columns can be easily added to a data frame
```{r}
mtcars$kpg <- mtcars$mpg * 8/5
mtcars[1:10,c("mpg","kpg")]
```

* The apply function can be used to process data frames to get the mean of a column
```{r}
apply(mtcars,2,mean)
```

* The sapply function can be used to process data frames to get the mean of a column
```{r}
sapply(mtcars,mean)
```

* The lapply function can be used to process data frames to get the mean of a column
```{r}
lapply(mtcars,mean)
```
* The package nycflights13 contains an excellent data set, this includes over 300,000 individual flight records
```{r}
summary(nycflights13::flights)
```

* A section of the data is shown here
```{r}
n <- sample(nrow(nycflights13::flights),10)
nycflights13::flights[n,c("year",
                          "month",
                          "day",
                          "dep_delay", 
                          "arr_delay", 
                          "origin", 
                          "dest", 
                          "distance")]
```




