---
output: github_document
title:  Lecture 6 (dplyr) - Summary
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
The dplyr package is a key part of the tidyverse in R, and provides a number of useful functions to process rectangular data in R. The main functions are the "five verbs", which are filter(), arrange(), select(), mutate() and summarise().

* **filter()** allows you to pick observations by their values, effectively filtering the available data set into a subset of rows that match certain conditions. We show the tibble mpg (contained in ggplot2), and then performa a filter operation on the model column (model=="a4")


```{r}
library(ggplot2)
library(dplyr)

mpg

filter(mpg,model=="a4")

```

* **arrange()** facilitates sorting of data based on a specified column, or number of columns. The default is ascending, and the function **desc()** can be used to sort in ascending order

```{r}

arrange(mpg,cty)

arrange(mpg,desc(cty))

```

* **arrange()** can also process multiple columns.

```{r}

arrange(mpg,desc(cty),desc(hwy))

```

* **select()** is used to reduce the number of columns in the tibble
```{r}

select(mpg,manufacturer,model,year,cty,hwy)

```


* **select()** is used to reduce the number of columns in the tibble
```{r}

select(mpg,manufacturer,model,year,cty,hwy)

```

* **select()** can work with a number of functions, for example ends_with() and contains()
```{r}

select(mpg,ends_with("r"))
select(mpg,contains("l"))

```


* **mutate()** creates new variables with functions of existing variables
```{r}
t <- mutate(mpg,CityHwayRatio=cty/hwy)
select(t,manufacturer,model, year, cty,hwy, CityHwayRatio)
```

* **summarise()** can collapse many values down to a single summary, using the function group_by()

```{r}
gr <- group_by(mpg,manufacturer)
summarise(gr,MinYear=min(year),MaxYear=max(year),AvrCity=mean(cty),AvrHwy=mean(hwy))
```

* **pull()** will return a column as a vector. By default, it returns the last column, otherwise the column number/name can be specified

```{r}
t <- filter(mpg,model=="a4")
t
pull(t,1)
```

* **case_when()** can perform a sequence of if-else commands, and can be used with mutate()
```{r}
t <- mutate(mpg,CarType=case_when(
    displ < 2.4 ~ "Small",
    displ < 4.6 ~ "Medium",
    TRUE        ~ "Large"
))
arrange(t,desc(displ))
```


* The **%>% (pipe operator)** can be used to create a workflow of chained operations.

```{r}
mpg %>% group_by(manufacturer) %>%
summarise(MinYear=min(year),
          MaxYear=max(year),
          AvrCity=mean(cty),
          AvrHwy=mean(hwy))
```