---
title: '10. Introduction to the S4 Object System'
author: "CT5102 - J. Duggan"
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(methods)
```

## S4 (Wickam 2019, Chapter 15)

- A more formal approach to functional OOP
- Underlying ideas similar to S3, but implementation is much stricter
- Makes use of specialised functions for:
  + Creating classes **setClass()**
  + Generics **setGeneric()**
  + Methods **setMethod()**
- Provides multiple inheritance and multiple dispatch
- A new component of S4 is the slot, a named component of an object that can be accessed using @ (pronounced at)
- Include **library(methods)** when using S4


## Basics

```{r,echo=T}
setClass("Person",
         slots = c(
           name = "character",
           age  = "numeric"
         )
)

john <- new("Person", name = "John Smith", age=35)
```

Given an S4 object, you can see its class with **is()** and access its slots with **\@** and **slot()**

## Calling S4 functions
```{r,echo=T}
is(john)
john@name
john@age
slot(john,"age")
slot(john,"name")
```

## Accessing slots: Guidelines
- Generally, only use @ in your methods
- Look for accessor functions that allow you to safely set and get slot values
- When you develop a class, provide your own access functions
- Creating a setter and getter for the age slot by
  + Creating getter and setter generics usiing **setGeneric()**
  + Defining methods with **setMethod()**

## getter for slot age
```{r,echo=T}
setGeneric("age", function(x)standardGeneric("age")) # getter

setMethod("age", "Person", function(x) x@age)

age(john)
```

## setter for slot age
```{r,echo=T}
setGeneric("age<-", function(x, value)standardGeneric("age<-")) # setter

setMethod("age<-", "Person", function(x, value){
  x@age <- value
  x
})
age(john) <- 29
age(john)

```

## Creating S4 Classes
- To define an S4 class, call setClass() with three arguments
  + The class **name**. By convention S4 class names use UpperCamelCase
  + A named character vector that describes the names and classes of the **slots** (fields). The pseudo-class ANY allows a slot to accept objects of any type
  + A **prototype**, a list of default values for each type (optional but should be provided)

## S4 class with 3 arguments
```{r,echo=T}
setClass("Person",
         slots = c(
           name = "character",
           age  = "numeric"
         ),
         prototype=list(
           name = NA_character_,
           age  = NA_real_
         )
)

test <- new("Person", name = "A.N. Other")
str(test)
```

## S4 class with inheritance
```{r,echo=T}
setClass("Employee",
         contains = "Person",
         slots = c(
           boss = "Person"
         ),
         prototype=list(
           boss = new("Person")
         )
)
str(new("Employee"))
```

## Helper
User facing classes should always be paired with a user-friendly helper. A helper should always:

- Have the same name as the class
- Have a thoughtfully crafted user interface with carefully chosen default values
- Create error messages tailored towards an end user
- Finish by calling **methods::new()**

## Example
```{r,echo=T}
Person <- function(name, age=NA){
  age <- as.double(age)
  new("Person", name=name, age=age)
}

p1 <- Person("A.N. Other")
p1
```

## Validator
- The constructor automatically checks that the slots have correct classes
- More complicated checks may be required
- For example, name and age may have to be the same vector length
- We can write a validator with **setValidity()**
```{r,echo=T,eval=F}
setValidity("Person", function(object) {
  if(length(object@name) != length(object@age)){
    "@name and @age must be the same length"
  }else{
    TRUE
  }
})
```


## Generics and Methods
- The job of a generic is to perform method dispatch (find the specific implementation for the combination of classes passed to the generic)
- To create a new S4 generic, call **setGeneric()** with a function that calls **stanardGeneric()**
- By convention, new S4 generics should use lowerCamelCase.
- It is bad practice to use {} as it triggers a special case that is more computationally expensive
```{r,echo=T,eval=F}
setGeneric("myGeneric", 
           function(x) standardGeneric("myGeneric"))
```

## Generics and Methods
- A generic isn't useful without some methods, and in S4 methods are defined with **setMethod()**
- There are three important arguments:
  + The name of the generic
  + The name of the class
  + The method itself
```{r,echo=T,eval=F}
setMethod("myGeneric", "Person", function(x){
  # method implementation
})
```

## Show method
- The most commonly defined S4 method that controls printing is **show()**.
- To define a method for an existing generic, you must first determine its arguments
- Our show method needs to have a single argument object
```{r,echo=T}
args(getGeneric("show"))
```

## Writing the method
```{r,echo=T}
setMethod("show", "Person", function (object){
  cat(is(object)[[1]], "\n",
  "   Name: ", object@name, "\n",
  "   Age:  ", object@age, "\n",
  sep="")
})

p1
```

## Accessors - getter function
- Slots should be considered an internal implementation detail: user code should avoid accessing them directly
- Typically, a generic will be defined so that multiple classes can use the same interface
```{r,echo=T}
setGeneric("name", function(x) standardGeneric("name"))
setMethod("name", "Person", function(x) x@name)

name(p1)
```

## Accessors - setter function
- If the slot is writeable, a setter function should be provided
- Should also include **validObject()** in the setter

```{r,echo=T}
setGeneric("name<-", function(x, value) 
            standardGeneric("name<-"))
setMethod("name<-", "Person", function(x, value){ 
  x@name <- value
  x})

name(p1) <- "Test Name"
name(p1)
```

## Exercises and Further features of S4
- Exercises
  + Add age() accessors for the Person class
  + In ther definition of the generic, why is it necessary to repeat the name of the generic twice
- Method Dispatch
  + Multiple inheritance - a class can have multiple parents
  + Multiple dispatch - a generic can use multiple arguments to pick a method


## Examples of S4 - Mapping functions
- The **sp** package defines a number of S4 classes for handling spatial data: points, lines and areas (Brunsdon and Comber 2019)
- The package **GISTools** contains useful data sets

|Without Attributes| With Attributes|ArcGIS equivalent |
|:------|:----------|:-----------|
|SpatialPoints|SpatialPointsDataFrame|Point shapefiles |
|SpatialLines| SpatialLinesDataFrame|Line Shapefiles|
|SpatialPolygons| SpatialPolygonsDataFrame|Polygon shapefiles|

## Georgia Data Set
```{r,echo=T,message=F}
library(GISTools)
library(sp)
library(tmap)

data(georgia)
class(georgia)
getSlots("SpatialPolygonsDataFrame")
```

## Exploring the data (1st county of 159)
```{r,echo=T,message=F}
length(georgia@polygons)
str(georgia@polygons[[1]])
```

## Getting the data
```{r,echo=T,message=F}
head(georgia@data)
```

## Displaying the Map
```{r,echo=T,message=F,fig.width=3, fig.height=2.5}
qtm(georgia,fill="red",style="natural")
```

## Filling by attribute
```{r,echo=T,message=F,fig.width=6, fig.height=2}
qtm(georgia,fill="MedInc",text="Name",text.size=0.1,
    format="World_wide",style="classic",
    text.root=5,fill.title="Median Income")
```


## New sf package
```{r,echo=T,message=F,fig.width=6, fig.height=2}
library(GISTools)
library(sp)
library(sf)
library(tmap)

data(georgia)
georgia_sf <- st_as_sf(georgia)
class(georgia_sf)
```

## New sf package
```{r,echo=T,message=F,fig.width=6, fig.height=2}
georgia_sf[1:2,]
```

## Displaying the Map (S3 class)
```{r,echo=T,message=F,fig.width=3, fig.height=2.5}
qtm(georgia_sf,fill="red",style="natural")
```

## Summary
- S4 is a more formal approach to functional OOP
- Underlying ideas similar to S3, but implementation is much stricter
- Include **library(methods)** when using S4
- Popular packages
  + **sp** for maps (**sf** uses **sp**)
  + Bioconductor (R libraries for BioInformatics)

```{r,echo=F,fig.width=2, fig.height=2}
# this is a hack to call plot but make the plot tiny
plot(1:1,axes=FALSE,xlab = "", ylab="",cex = .001)
```




