---
output: github_document
title:  Lecture 8 (stringr) - Summary
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
The package **stringr** supports string processing in R. A key idea is the use of regular expressions to search for terms. Text mining is also a fascinating area, and R offers further tools, for example see: [Text Mining with R)](http://tidytextmining.com/index.html), wriiten by Julia Silge and David Robinson.


Useful functions are

* **str_detect()**, which determines if a character vector matches a pattern, and returns a logical vector of the same length.


```{r}
library(stringr)
x <- c("Apple","Banana","Pear")
str_detect(x,"an")
x[str_detect(x,"an")]
```

* More complex string patterns can be formed using regular expressions
    - To match **.**, we need to use the escape sequence **\\** before the character **.**
    
    ```{r}
    x <- c("apple.com","Banana","Pear")
    x[str_detect(x,"e\\.com")]
    ```


    - Anchors can be used to match from the start (^) of a string, or from the end ($).

    
    ```{r}
    x <- c("Apple","Banana","Pear")
    x[str_detect(x,"^P")]
    x[str_detect(x,"a$")]
    ```


    - **\\d**  – matches any digit
    - **\\s** – matches any whitespace
    - **[abc]** – matches a, b or c
    - **[^abc]** – matches anything except a, b or c.
    
    
    ```{r}
    x <- c("Dublin","Galway","Cork 1","DCDCDCdcdcdc")
    x[str_detect(x,"\\d")]
    x[str_detect(x,"\\s")]
    x[str_detect(x,"[DC]")]
    ```
    
    
    - For repetition, ? for (0 or 1) * for (0 or more), + for (1 or more)

    ```{r}
    x <- c("Italy","Germany","Greece")
    x[str_detect(x,"e?")]
    x[str_detect(x,"e*")]
    x[str_detect(x,"e+")]
    ```

    - {n}: exactly n
    - {n,}: n or more
    - {n,m}: between n and m
    
    ```{r}
    x <- c("Italy","Germany","Greece")
    x[str_detect(x,"e{2}")]
    x[str_detect(x,"e{1,}")]
    x[str_detect(x,"e{2,2}")]
    ```
    
   
* **str_subset()**, can be used to filter the data set based on a regular expression

```{r}
s <- sentences[1:5]
str_subset(s,"Rice|blue")
```


* Grouped matches can be used to extract parts of a complex match

```{r}
s <- sentences[1:5]
noun <- "(a|the) ([^ ]+)"
str_subset(s,noun)

str_subset(s,noun) %>% str_extract(noun)

```

