---
output: github_document
title:  Lecture 2 (Lists and Functions) - Summary
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
Here are some of the key points from the lecture. See also the sample code solutions in the code folder.

* Unlike atomic vectors, lists can contain many different types (incuding lists)

```{r}
l <- list(1:3, "Test",c(T,F,T,T))
str(l)
```

* Lists can be subsetted like an atomic vector. There are two ways to subset, preserving [] and simplifying [[]]

```{r}
l <- list(1:3, list(2:3,4:5))
str(l)
# this preserves the list structure
l1 <- l[1]
str(l1)

# This simplifies (and returns the contents of the first element)
c1 <- l[[1]]
str(c1)

# This finds the second vector that is part of the 2nd list element
c2 <- l[[2]][[2]]
str(c2)

# This finds the first element of the first vector that is part of the 2nd list element
c3 <- l[[2]][[1]][1]
str(c3)
```

* Here is a nice example of how a list is used in machine learning/data science
```{r}
mod <- lm(data=datasets::faithful,eruptions~waiting)
str(mod)
summary(mod)
```

* Functions are defined using the keyword function, and return the last evaluated expression

```{r}
evens <- function (v){
  v[v %% 2 == 0]
}

evens(1:10)
```

* Functions can be passed to functions
```{r}
t1 <- function (f, v){
  f(v)
}

t1(min,1:10)
t1(max,1:10)
t1(sum,1:10)
```

* Functions can have default values
```{r}
evens <- function (v, dec=F){
  sort(v[v %% 2 == 0],decreasing=dec)
}

evens(1:10)
evens(1:10, dec = T)
```

* When functions are created, they have access to all the variables defined in that environment
```{r}
c<-10
test <- function (a, b){
  a+b+c
}

test(10,20)
```

* The environment in R is a hierarchy, with the Global Environment at the bottom of this structure.
```{r}
search()
```

* You can find the parent of any environment
```{r}
parent.env(globalenv())
```

* R will find the first occurance of a variable/function in the hierachy. For example
```{r}
environment(mean)
x <- mean(1:10)
x
```


* You can create a function with the same name as another function (in a different package) in your Global Envirnoment. This could lead to tricky situations.
```{r}
mean<-function(v)max(v)
environment(mean)
x <- mean(1:10)
x
```
* You can also ensure that you call the correct function by prefixing the package name and using ::
```{r}
mean<-function(v)max(v)
environment(mean)
x <- base::mean(1:10)
x
```
