---
output: github_document
title:  Lecture 1 (Atomic Vectors) - Summary
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
Here are some of the key points from the lecture. See also the sample code solutions in the code folder.

* Atomic vectors have only one data type.

```{r}
v1 <- 1:5
str(v1)

v2 <- c(T,F)
str(v2)
```

* When combining different data types in a vector, R will coerce to the most flexible data type

```{r}
v3 <- c(v1, v2)
str(v3)
```

* Functions covered in the lecture include **c()**, **rep()**, **str()**, **min()**, **max()**, **median()**, **mean()**, **names()**

* R supports vectorisation, whereby the same function can be applied to each element. For example:
```{r}
x1 <- (1:5)^2
y1 <- sqrt(x1)
print(y1)
```

* Filtering a vector starts with an index value of 1. 
```{r}
u1 <- 1:10
y1 <- u1[1]
print(y1)
y2 <- u1[1:2]
print(y2)

```

* A vector can be filtered on the element name also
```{r}
u2 <- 1:5
names(u2) <- c("p1","p2","p3","p4","p5")
print(u2)
print(u2["p3"])

```


* A logical vector can be used to filter a vector. For example, to print every second element of a vector, the following can be used:
```{r}
x1 <- (1:10)^2
y1 <- x1[c(T,F)]
print(y1)
```

* The function **which()** can find the index for a value
```{r}
x1 <- (1:10)^2
loc <- which(x1 == 81)
print(x1)
print(loc)
print(x1[which(x1==81)])
```
